---
title: | 
  | Econometrics III 
  | Assignment Part II
author: | 
  | Thao Le (523716) 
  | David Gyaraki (582340)
format: pdf
pdf:
  documentclass: article
  cite-method: biblatex
editor: source
include-in-header:
  text: |
    \addtokomafont{disposition}{\rmfamily}
    \usepackage{amsmath}
    \newcommand{\bm}{\symbf}
    \newcommand{\T}{\text{T}}
    \newcommand{\pl}{\text{plim}}
    \newcommand{\brefsection}[1]{Section \textcolor{blue}{\ref{#1}}}
    \newcommand{\beqref}[1]{Equation \textcolor{blue}{\eqref{#1}}}
    \newcommand{\breftable}[1]{Table \textcolor{blue}{\ref{#1}}}
    \newcommand{\breffig}[1]{Figure \textcolor{blue}{\ref{#1}}}
    \usepackage{fancyvrb}
    \usepackage{dcolumn}
    \usepackage{fvextra}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{
      showspaces = false,
      showtabs = false,
      breaklines,
      commandchars=\\\{\}
    }
pdf-engine: xelatex
cap-location: top
toc: true
toc-title: Contents
number-sections: true
mainfont: Arial
setspace:
  linestretch: 1.25
fig-align: center
table-align: center
fig-pos: H
table-pos: H
execute:
  echo: true
  warning: false
  eval: true
code-line-numbers: false
colorlinks: true
code-block-bg: darkgray
df-print: default
highlight-style: arrow-dark
biblio-title: References
---

\clearpage

```{r setup part1, echo=TRUE, results='hide'}
# load packages
if(!require(pacman)){install.packages("pacman")}

p_load(devtools,tidyverse,dplyr,ggplot2,latex2exp,stargazer, fixest, modelsummary, knitr, readr, tseries, lmtest, forecast, dynlm)

dfAssign_p2 <- as.data.frame(read_csv("data/data_assign_p2.csv"))

# Encode quarters
dfAssign_p2 <- cbind(dfAssign_p2,c(seq(1,nrow(dfAssign_p2),length.out=nrow(dfAssign_p2))))
colnames(dfAssign_p2) <- c("obs", "GDP_QGR", "UN_RATE", "ind")
```

\section{Question 1} \label{q1}

```{r q1}
ggplot(dfAssign_p2) +
  geom_point(aes(x = ind, y = GDP_QGR, color="GDP growth")) + 
  geom_line(aes(x = ind, y = GDP_QGR, color="GDP growth")) +
  geom_point(data=dfAssign_p2, aes(x=ind, y=UN_RATE, color="Unemployment"), shape=3) + 
  ylab("GDP growth and unemployment rate in %") + 
  xlab("Quarters since 1987 Q2") +
  scale_color_manual(values = c(
    "Unemployment" = "red",
    "GDP growth" = "blue"))

plot.ts(dfAssign_p2[,2:3], plot.type = c("multiple"), main="GDP quarterly growth and unemployment rate since 1987 Q2")
```

MAYBE SOME TEXT HERE? WE HAD NO QUESTIONS BUT JUST TO DESCRIBE THE PLOT?

```{r}
# Creating the autoregression of GDP growth
ar4_gdp <- arima(dfAssign_p2$GDP_QGR, order=c(4,0,0))
coeftest(ar4_gdp)

ar1_3_4_gdp <- arima(dfAssign_p2$GDP_QGR, order=c(4,0,0), fixed = c(NA,0,NA,NA,NA))
coeftest(ar1_3_4_gdp)

ar1_3_gdp <- arima(dfAssign_p2$GDP_QGR, order=c(3,0,0), fixed = c(NA,0,NA,NA))
coeftest(ar1_3_gdp)
# Stop here, 3rd lag coefficient is significant

dfAssign_p2$GDP_QGR <- ts(dfAssign_p2$GDP_QGR, start = c(1987, 2), frequency = 4)
dfAssign_p2$UN_RATE <- ts(dfAssign_p2$UN_RATE, start = c(1987, 2), frequency = 4)

adl_4_4 <- dynlm(UN_RATE ~ L(UN_RATE, 1) + L(UN_RATE, 2) + L(UN_RATE, 3) + L(UN_RATE, 4) + L(GDP_QGR, 1) + L(GDP_QGR, 2) + L(GDP_QGR, 3) + L(GDP_QGR, 4), data = dfAssign_p2)
  
coeftest(adl_4_4)

adl_3_3 <- dynlm(UN_RATE ~ L(UN_RATE, 1) + L(UN_RATE, 2) + L(UN_RATE, 3) + L(GDP_QGR, 1) + L(GDP_QGR, 2) + L(GDP_QGR, 3), data = dfAssign_p2)
  
coeftest(adl_3_3)

adl_3_2 <- dynlm(UN_RATE ~ L(UN_RATE, 1) + L(UN_RATE, 2) + L(UN_RATE, 3) + L(GDP_QGR, 1) + L(GDP_QGR, 2), data = dfAssign_p2)
  
coeftest(adl_3_2)

adl_3_1 <- dynlm(UN_RATE ~ L(UN_RATE, 1) + L(UN_RATE, 2) + L(UN_RATE, 3) + L(GDP_QGR, 1), data = dfAssign_p2)
  
coeftest(adl_3_1)
```

IT IS YET UNCLEAR HOW TO PROGRESS WITH THE RESTRICTIONS. DOES ONE RESTRICT CO-JOINTLY, START WITH X OR START WITH Y ALONE?

\section{Question 2} \label{q2}

```{r q2}

```




\section{Question 3} \label{q3}

```{r q3}

```



\section{Question 4} \label{q4}

```{r q4}

```



\section{Question 5} \label{q5}

```{r q5}

```



\section{Question 6} \label{q6}

\subsection{Question 6a} \label{q6a}


\subsection{Question 6b} \label{q6b}





